Statistical Machine Learning Models on A Bank Marketing Dataset

### 2025-12-09 Slides.Rmd
final slides title without numbers

## KNN 1/3: Feature Preparation 
* **KNN** requires scaling: all numeric predictors were **standardized** (mean=0, SD=1) to prevent features with larger scales from dominating the distance calculation.
* The non-parametric nature means **no explicit feature selection** via coefficients, but the scaling step is vital for ensuring equal contribution from all dimensions.

@@

## KNN 2/3: Hyperparameter Tuning 
* The sole hyperparameter, **$k$** (the number of neighbors), determines the model's complexity (i.e., the bias-variance trade-off).
* **5-Fold Cross-Validation** was performed across a range of $k$ values to find the optimal $k$ that minimized the mean CV error.

@@

## KNN 3/3: Evaluation & Test Performance
* The final KNN model (with optimal $k$) was applied to the **Test Set**. We evaluate performance using **AUC** and compare it to the LogReg models.
* KNN generally provides a good intermediate balance—often capturing non-linearity better than LogReg but being less computationally expensive than boosting.

@@

## Boosting 1/3: Model Formulation

@@

## Boosting 2/3: Hyperparameter Tuning 
- We tuned the number of trees using OOB error estimation to prevent overfitting.
- The OOB method selected 1100 trees as optimal.
- This choice balances model complexity and generalization.

@@

## Boosting 3/3: Evaluation & Importance 
- The final Boosting model achieved the highest AUC and sensitivity among all models tested.
- Variable Importance Plot shows which features influence predictions most.
- The importance rankings align with key predictors identified by Logistic Regression.

>>>>

## KNN: Feature Preparation 
* **KNN** requires scaling: all numeric predictors were **standardized** (mean=0, SD=1) to prevent features with larger scales from dominating the distance calculation.
* The non-parametric nature means **no explicit feature selection** via coefficients, but the scaling step is vital for ensuring equal contribution from all dimensions.

@@

## KNN: Hyperparameter Tuning 
* The sole hyperparameter, **$k$** (the number of neighbors), determines the model's complexity (i.e., the bias-variance trade-off).
* **5-Fold Cross-Validation** was performed across a range of $k$ values to find the optimal $k$ that minimized the mean CV error.

@@

## KNN: Evaluation & Test Performance
* The final KNN model (with optimal $k$) was applied to the **Test Set**. We evaluate performance using **AUC** and compare it to the LogReg models.
* KNN generally provides a good intermediate balance—often capturing non-linearity better than LogReg but being less computationally expensive than boosting.

@@

## Boosting: Model Formulation

@@

## Boosting: Hyperparameter Tuning 
- We tuned the number of trees using OOB error estimation to prevent overfitting.
- The OOB method selected 1100 trees as optimal.
- This choice balances model complexity and generalization.

@@

## Boosting: Evaluation & Importance 
- The final Boosting model achieved the highest AUC and sensitivity among all models tested.
- Variable Importance Plot shows which features influence predictions most.
- The importance rankings align with key predictors identified by Logistic Regression.
